\documentclass[10pt,a4paper]{article}

\usepackage[margin=1.6cm]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=blue!50!black
}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.45em}
\setlist[itemize]{leftmargin=*, itemsep=0.2em, topsep=0.2em}
\setlist[enumerate]{leftmargin=*, itemsep=0.2em, topsep=0.2em}

\title{\vspace{-1.2em}\textbf{FPGA Wand: A Multi-Node Interactive Drawing System with Local FPGA Processing and Cloud Coordination}\vspace{-0.8em}}
\author{[Author Name]}
\date{}

\begin{document}
\maketitle
\vspace{-1.4em}

\section{Purpose of the System}

The aim of this project was to build a distributed interactive system in which
multiple FPGA SoC nodes perform meaningful local processing, exchange compact
events with a cloud service, and can also be influenced by server-side control.
The final implemented system is an interactive ``wand drawing'' platform built
around PYNQ boards, a camera-based local processing pipeline, and a cloud
service called \emph{Wand Brain}. Each node observes a bright LED wand through a
camera, reduces each frame locally to a centroid-like point, and transmits this
point stream to the cloud as stroke events. The cloud reconstructs attempts,
renders them live in a browser dashboard, scores finalized strokes against
target templates, maintains leaderboards, and sends control settings back to
the nodes.

This final system was chosen because it satisfies all minimum functional
requirements while remaining technically rich. It demonstrates local processing
on the FPGA SoC device, server-side event processing, bidirectional
communication between node and server, and concurrent operation of at least two
nodes. The design also leaves clear space for extension, including alternative
templates, richer scoring, and more advanced node-side vision.

\section{Overall Architecture}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  node distance=0.85cm and 0.65cm,
  box/.style={draw, rounded corners, align=center, minimum width=2.4cm, minimum height=0.9cm, fill=blue!4},
  smallbox/.style={draw, rounded corners, align=center, minimum width=2.15cm, minimum height=0.8cm, fill=green!4},
  cloudbox/.style={draw, rounded corners, align=center, minimum width=2.65cm, minimum height=0.95cm, fill=orange!6},
  line/.style={-{Latex[length=2mm]}, thick}
]
\node[box] (camera) {USB Camera\\BGR Frames};
\node[box, right=of camera] (ps) {PYNQ PS\\OpenCV preprocessing\\UDP/HTTP logic};
\node[smallbox, below=of ps] (dma) {AXI DMA};
\node[smallbox, left=of dma] (pl) {Custom PL IP\\frame\_centroid\_axi};
\node[cloudbox, right=of ps, xshift=1.0cm] (brain) {EC2 ``Wand Brain''\\UDP ingest, render, score, persist};
\node[cloudbox, below=of brain] (db) {SQLite history\\leaderboards};
\node[box, above=of brain] (web) {Browser dashboard\\live view and control};

\draw[line] (camera) -- node[above, font=\scriptsize] {frame capture} (ps);
\draw[line] (ps) -- node[right, font=\scriptsize] {binary frame buffer} (dma);
\draw[line] (dma) -- node[below, font=\scriptsize] {AXI stream} (pl);
\draw[line] (pl) -- node[left, font=\scriptsize] {AXI-Lite stats} (ps);
\draw[line] (ps) -- node[above, font=\scriptsize] {24-byte UDP point packets} (brain);
\draw[line] (brain) -- node[right, font=\scriptsize] {finalized attempts} (db);
\draw[line] (web) -- node[right, font=\scriptsize] {HTTP polling} (brain);
\draw[line] (brain.west) to[bend right=18] node[below, font=\scriptsize] {HTTP node control} (ps.east);
\end{tikzpicture}
\caption{End-to-end architecture of the final system. The fast data plane uses
UDP point events from the node to the cloud, while the control plane uses HTTP
from the cloud back to the node.}
\label{fig:system-architecture}
\end{figure}

Figure~\ref{fig:system-architecture} shows the split between the local node and
the cloud service. The local node combines the programmable logic (PL) and the
processing system (PS) of the PYNQ board. The PS is responsible for camera
capture, grayscale conversion, resizing to $640 \times 480$, thresholding,
runtime filtering, stroke management, and networking. The PL accelerates the
repetitive per-pixel reduction stage by scanning the thresholded binary frame
and exposing compact image statistics instead of a full processed image.

\subsection*{Node-side PS/PL partition}

The adopted FPGA design consists of a custom AXI-stream IP,
\texttt{frame\_centroid\_axi}, attached to an AXI DMA engine and accessed by the
PS through AXI-Lite. After the PS thresholds a frame, it writes the binary
image into a DMA buffer in memory. The DMA then streams the image into the PL,
where four pixels are processed per 32-bit beat. The custom IP accumulates:
\begin{itemize}
  \item the horizontal coordinate sum $\sum x$,
  \item the vertical coordinate sum $\sum y$,
  \item the foreground-pixel count,
  \item a frame identifier, and
  \item a valid bit.
\end{itemize}
The PS reads these registers through MMIO and computes the centroid using
$c_x=\sum x / \mathrm{count}$ and $c_y=\sum y / \mathrm{count}$. This is a
deliberate partition: the PL performs the regular streaming arithmetic, while
the PS keeps the flexible logic in software, including filtering, control
handling, sketch rendering, and networking.

\subsection*{Cloud-side runtime}

The cloud service is a FastAPI application deployed on EC2. It is divided into
two layers. The base runtime receives UDP packets on port 41000, validates and
parses them into structured point events, groups them into live attempts, and
renders both live and finalized drawings. A wrapper layer adds persistence,
template selection, node control, recent-attempt views, leaderboards, and the
HTML dashboard. The server also exposes health, template, and debugging routes
over HTTP on port 8000.

The browser frontend is a deliberately simple single-page console. It polls the
backend rather than using WebSockets, which made the system easier to debug and
deploy during rapid iteration. Live frame polling is configured at 180\,ms,
selected-wand status at 500\,ms, and health/leaderboard polling at slower
intervals. This produces a responsive operator view while keeping the client
logic lightweight.

\section{Performance Metrics and Their Rationale}

Because the project is an interactive distributed system, its performance
cannot be described by a single number. The most relevant metrics are the ones
that determine whether a user can draw reliably, whether the cloud reconstructs
the intended stroke, and whether the FPGA acceleration is justified.

\begin{table}[t]
\centering
\small
\begin{tabularx}{\linewidth}{>{\raggedright\arraybackslash}p{2.8cm} >{\raggedright\arraybackslash}p{4.1cm} X}
\toprule
\textbf{Metric} & \textbf{Why it matters} & \textbf{How it is evidenced in this project} \\
\midrule
End-to-end stroke correctness & The system must turn a physical wand motion into the correct live and finalized attempt. & Verified through live rendering, finalized PNG endpoints, scoring output, and synthetic plus live-board tests. \\
Interaction responsiveness & Drawing is only usable if the system responds quickly enough for human feedback. & Compact 24-byte point packets, local point gating on the node, and a responsive polling dashboard were used to reduce perceived delay. \\
Multi-node isolation & Minimum requirements demand concurrent receive/transmit from at least two nodes. & \texttt{device\_number} and \texttt{wand\_id} are carried through transport, backend state, and UI; two-node live validation was completed. \\
Control-plane correctness & The server must be able to affect local processing on the nodes. & HTTP control and acknowledgement routes were added, with runtime changes such as \texttt{enabled}, \texttt{armed}, thresholds, and recalibration. \\
FPGA efficiency & The local hardware must provide useful acceleration without becoming too large or fragile to integrate. & The final PL design performs only the image-reduction stage in hardware, streams four pixels per beat, and met timing closure. \\
\bottomrule
\end{tabularx}
\caption{Chosen performance metrics for the implemented system.}
\label{tab:metrics}
\end{table}

The system was intentionally evaluated in terms of correctness, responsiveness,
and deployability rather than headline throughput alone. For example, the UDP
protocol sends one point per 24-byte datagram rather than a frame or a large
batch. This keeps the transport lightweight and suits interactive drawing, for
which occasional packet loss is less harmful than retransmission delay. By the
same logic, the frontend uses frequent polling only where it most affects user
feedback, such as live frame refresh, while slower-changing views use longer
poll intervals.

\section{Design Decisions and Evaluation}

\subsection*{Hardware/software split}

One of the most important design decisions was to avoid placing the entire
vision pipeline into the PL. Instead, the PL implements the streaming reduction
stage only, while the PS performs thresholding, centroid division, validity
rules, and network communication. This trades maximum hardware throughput for
integration reliability and flexibility. In practice, it was a good compromise:
the custom IP still accelerates the repetitive per-pixel work, but changes to
tracking behaviour, stroke rules, and cloud interaction remain easy to make in
software.

\subsection*{UDP data plane, HTTP control plane}

The system uses two protocols with distinct roles. The data plane uses the
\texttt{wb-point-v1} UDP packet format, where each datagram contains one point,
stroke flags, node identity, and a timestamp. This design keeps latency low and
allows explicit stroke lifecycle events such as \texttt{STROKE\_START} and
\texttt{STROKE\_END}. In contrast, the control plane uses HTTP because control
messages need versioning, readability, and easy debugging. This made it
straightforward to add node control, health checks, and acknowledgements
without destabilizing the point-stream path.

\subsection*{In-memory live state with persisted finalized history}

The cloud keeps active strokes in memory and only persists attempts once they
are finalized. This was another deliberate trade-off. It avoids database
overhead on every incoming point and allows the backend to update live images
very cheaply. The disadvantage is that active in-memory state is lost if the
server restarts. For the current project stage this was acceptable, because
historical scoring, recent attempts, and leaderboards all depend on finalized
rather than partially complete strokes.

\subsection*{Single-page dashboard}

The dashboard is implemented as a single HTML page with embedded JavaScript.
This choice reduced framework overhead and kept the live demo easy to deploy on
EC2. It also supported rapid additions such as node control, stroke timers, and
leaderboards. The downside is that the page relies on polling rather than true
server push, so it is better suited to this lab-scale console than to a large
production deployment. For the project brief, however, the simplicity and
clarity of this approach were strong advantages.

\section{Testing Approach and Validation}

\begin{figure}[t]
\centering
\begin{tikzpicture}[
  node distance=0.55cm and 0.6cm,
  flowstep/.style={draw, rounded corners, align=center, minimum width=2.6cm, minimum height=0.85cm, fill=gray!8},
  flow/.style={-{Latex[length=2mm]}, thick}
]
\node[flowstep] (cloud) {Cloud sanity check\\health, templates, dashboard};
\node[flowstep, right=of cloud] (udp) {UDP connectivity\\synthetic stroke injection};
\node[flowstep, right=of udp] (smoke) {Smoke test\\render + finalize paths};
\node[flowstep, below=of udp] (board) {PYNQ validation\\camera, DMA, MMIO, control};
\node[flowstep, left=of board] (multi) {Two-node test\\simultaneous draw paths};
\node[flowstep, right=of board] (ui) {Frontend regression\\score, timer, leaderboard};

\draw[flow] (cloud) -- (udp);
\draw[flow] (udp) -- (smoke);
\draw[flow] (smoke) -- (ui);
\draw[flow] (smoke) -- (board);
\draw[flow] (board) -- (multi);
\draw[flow] (board) -- (ui);
\end{tikzpicture}
\caption{Testing flow used to move from isolated service checks to full
integrated multi-node validation.}
\label{fig:test-flow}
\end{figure}

The testing strategy combined tool-based validation with live integrated
demonstrations. Figure~\ref{fig:test-flow} summarizes the progression from
isolated checks to full-system tests.

At the cloud-only level, the backend was first validated through a health
check, template enumeration, and dashboard loading. Lightweight Python tools
were then used to inject synthetic UDP strokes and verify that the latest
attempt, live image, and finalized image endpoints behaved correctly. A further
smoke test exercised the full finalize-and-render path without requiring the
camera or FPGA board, which made regression testing much faster.

Board-level validation focused on the FPGA SoC runtime. This included camera
bring-up, repeated DMA/MMIO frame updates, correct extraction of centroid
statistics, and successful communication with the cloud over both UDP and HTTP.
The control-plane path was checked by observing control acknowledgements under
the node-control API. These tests confirmed that the server could affect local
node behaviour by enabling or disabling tracking, changing thresholds and gap
timeouts, and triggering recalibration.

The final integration stage used real hardware, the actual FPGA design, the EC2
deployment, and the browser dashboard. Both one-node and two-node sessions were
completed. The following final-system behaviours were validated:
\begin{itemize}
  \item correct UDP receive and parse on the cloud;
  \item live stroke plotting during active drawing;
  \item finalized-attempt rendering and template scoring;
  \item recent-attempt and database-backed history updates;
  \item HTML-based node control and acknowledgement;
  \item leaderboard promotion and champion-name claim logic; and
  \item simultaneous receive/transmit from two nodes.
\end{itemize}

This staged approach was particularly important because several integration
issues appeared only during full-system testing. Examples included camera
bring-up fragility on the PYNQ board, the need to separate backend-owned
\texttt{attempt\_id} values from incoming UDP \texttt{stroke\_id} values, and
ordering the recent-attempt list by finalization time rather than creation time.
Each issue was addressed without changing the overall architecture, which is a
useful indication that the design boundaries were sensible.

\section{Device Resources and Hardware Implementation}

The adopted FPGA implementation is intentionally compact. Rather than storing
frames or implementing a full image-processing graph in hardware, the custom PL
block performs only what is both repetitive and structurally regular: scanning
the thresholded image stream and accumulating foreground statistics. This has
several resource-related advantages:
\begin{itemize}
  \item no connected-component analysis in hardware;
  \item no hardware frame buffer for the processed image;
  \item no hardware division for centroid calculation;
  \item a narrow AXI-Lite register interface for result readback; and
  \item reuse of the PS for all policy, filtering, and network logic.
\end{itemize}

This means the PL resource footprint is dominated by simple counters,
accumulators, stream handling, and interface logic rather than by large memory
structures or complex arithmetic. Architecturally, this is a strong fit for the
project: the node still benefits from genuine hardware acceleration, but the
design remains understandable and practical to debug.

The archived implementation evidence includes a timing summary rather than a
complete post-implementation utilization table. Even so, the available timing
result is encouraging: the implemented design closed timing with setup worst
negative slack of 11.636\,ns, hold slack of 0.045\,ns, pulse-width slack of
8.750\,ns, and zero failing endpoints. This confirms that the final chosen PL
path was not only functionally useful but also straightforward to implement
reliably on the target platform.

\section{Conclusion}

The final system demonstrates a complete multi-node FPGA-to-cloud interactive
application rather than a collection of isolated components. Each node performs
local camera processing on the PYNQ board, the PL accelerates the image
reduction stage, the PS packages and transmits stroke events over UDP, the
cloud reconstructs and scores attempts, and the server can also influence the
node through HTTP control. The system therefore satisfies the key coursework
requirements while also providing additional functionality such as live plotting,
template-based scoring, stroke timers, and leaderboards.

From an engineering perspective, the most successful design choice was the
clear separation of responsibilities: regular streaming arithmetic in the PL,
flexible policy and networking in the PS, low-latency event transport over UDP,
and coordination, persistence, and presentation in the cloud service. The main
remaining limitations are also clear. The current centroid-based tracker cannot
separate multiple bright blobs, active live state is not preserved across a
server restart, and the dashboard relies on polling rather than push updates.
These limitations provide a direct roadmap for future work, but they do not
prevent the present system from operating as a coherent and demonstrably working
multi-node information-processing platform.

\end{document}
